Evolution of pre-copulatory and post-copulatory inbreeding behaviour: a computational

### 2016-09-16 ms_pre_post.Rmd
Finish a rough draft for review among colleagues

---
title: 'Evolution of pre-copulatory and post-copulatory inbreeding behaviour: a computational
  approach'
author: "A. Bradley Duthie et al."
date: "August 24, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

@@

Specific mechanisms of post-copulatory inbreeding avoidance are difficult to identify, but recent experimental studies have shown how females can successfully bias fetilisation after mating has occurred in both vertebrates [e.g., @Gasparini2011] and invertebrates [e.g., @Tuni2013]. For example, @Gasparini2011 artificially inseminated female guppies (*Poecilia reticulata*) with equal quantities of sperm from full-siblings and unrelated males, finding that fertilisation success was higher in non-relatives because the velocity of full siblings' sperm was reduced by females' ovarian fluids. Biased fertilisation toward non-relatives might also be accomplished through selective storage of sperm. For example, in black field crickets (*Teleogryllus commodus*), females attempt to remove the spermataphores of unwanted male mates after copulation, while male mates guard females to prevent spermatophore removal [@Bussiere2006]. Even if females are unable to remove spermatophores of unwanted mates, @Tuni2013 found evidence that female *T. commodus* are capable of controlling the transfer of sperm to their spermatheca, most likely through physical control of the spermathical duct, and can reduce sperm storage despite being guarded by a related male. Our results suggest that evolution of such post-copulatory inbreeding avoidance might increase selection for polyandry, especially if pre-precopulatory inbreeding avoidance is costly (Figure 1) or constrained (Figure 3).

@@

@@

Studies of multiple species have found evidence for female pre-copulatory inbreeding avoidance [e.g., @Lihoreau2007; @Kuriwada2011; @Kingma2013], and for post-copulatory inbreeding avoidance [e.g., @Simmons2006; @Bretman2009]. Other studies have

Few studies address evolution of both pre-copulatory and post-copulatory inbreeding strategies, but existing empirical studies are consistent with our model predictions. @Liu2014 found evidence for pre-copulatory, but not post-copulatory inbreeding avoidance in an experimental study of cabbage beetles (*Colaphellus bowringi*). In our model, post-copulatory

However, empirical hypothesis testing typically does not attempt to predict *a priori* whether or not inbreeding avoidance will be enacted before or after copulation, or if both pre-copulatory and post-copulatory inbreeding avoidance will evolve. Here we have shown that evolution of inbreeding strategies and associated polyandry are predicted to affected by direct costs, and constraints and adaptations of

@@

@@

\vspace{150 mm}

>>>>

---
title: 'Evolution of pre-copulatory and post-copulatory inbreeding behaviour: a computational
  approach'
author: "A. Bradley Duthie et al."
date: "August 24, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 2
csl: evolution.csl
bibliography: ms_pre_post.bib
---

<!--- Need to do 40k generations to see allele values level off, 40 replicates each:

Figure    1: 100 hours total
Figures   2: 48  hours total
Figures   3: 48  hours total
Figures   4: 48  hours total
Figures   5: 48  hours total

Nah, these are too much. I've decided to increase to 40k generations for the costs only (100 hours total). Reasoning for this is that when we look at the fixed traits, we are really interested in whether or not the initial invasion changes, hence fewer generations are better. When we start everything at zero and let it evolve, then long-term stuff becomes more important, so 40k are useful.

--->

@@

<!--- Some of these conclusions will need to be changed in light of the need for more generations of data: Strategies were concluded to be non-negligible, they they actually become negligible given a sufficient amount of time. --->

@@

Specific mechanisms of post-copulatory inbreeding avoidance are difficult to identify, but recent experimental studies have shown how females can successfully bias fetilisation after mating has occurred in both vertebrates [e.g., @Gasparini2011] and invertebrates [e.g., @Tuni2013]. For example, @Gasparini2011 artificially inseminated female guppies (*Poecilia reticulata*) with equal quantities of sperm from full-siblings and unrelated males, finding that fertilisation success was higher in non-relatives because the velocity of full siblings' sperm was reduced by females' ovarian fluids. Biased fertilisation toward non-relatives might also be accomplished through selective storage of sperm. For example, in black field crickets (*Teleogryllus commodus*), females attempt to remove the spermataphores of unwanted male mates after copulation, while male mates guard females to prevent spermatophore removal [@Bussiere2006]. Even if females are unable to remove spermatophores of unwanted mates, @Tuni2013 found that female *T. commodus* are capable of controlling the transfer of sperm to their spermatheca, most likely through physical control of the spermathecal duct, and can reduce sperm storage despite being guarded by a related male. Our results suggest that evolution of such post-copulatory inbreeding avoidance might increase selection for polyandry, especially if pre-copulatory inbreeding avoidance is costly (Figure 1) or constrained (Figure 3).

@@

Constraints on pre-copulatory and post-copulatory inbreeding avoidance might vary widely in empirical systems. Female mate choice might be evolutionarily constrained if females are unable to discriminate among kin and therefore incapable of choosing to avoid inbreeding [Duthie and Reid *in press*; @Penn2010]. Similarly, females might choose to avoid inbreeding, but be ecologically constrained in their choice by the availability of males if few non-relatives are available [e.g., @Kokko2006a; @Kokko2013], or if females are unable to resist male coercion given sexual conflict [@Arnqvist2005, pp. 44-91]. Females might also be unable to discriminate between sperm of relatives versus non-relatives, or be unable to successfully bias fertilisation toward non-relative's sperm. Our results suggest that such constraints on one strategy of inbreeding avoidance (pre-copulatory versus post-copulatory) will strengthen selection on a complementary strategy, perhaps making inbreeding avoidance more detectable in empirical systems.

@@

Studies of multiple species have found evidence for female pre-copulatory inbreeding avoidance [e.g., @Lihoreau2007; @Kuriwada2011; @Kingma2013; @Fischer2015], and for post-copulatory inbreeding avoidance [e.g., @Simmons2006; @Bretman2009]. Further, evidence of both pre-copulatory and post-copulatory inbreeding avoidance has been found for a handful of well-studied species, including Trinidadian guppies [*Poecilia reticulata*; @Gasparini2011; Daniel2015] and house mice [*Mus domesticus*; @Potts1991; @Firman2015], while @Liu2014 found evidence for pre-copulatory, but not post-copulatory, inbreeding avoidance in an experimental study of cabbage beetles (*Colaphellus bowringi*). These empirical studies demonstrate that evolution of all combinations of pre-copulatory and post-copulatory inbreeding avoidance are possible, but no general modelling attempts to predict *a priori* whether or not inbreeding avoidance will be enacted before or after copulation, or if both pre-copulatory and post-copulatory inbreeding avoidance will evolve. Here we have shown that evolution of inbreeding strategies and associated polyandry are predicted to be affected by direct costs, and constraints and adaptations of complementary inbreeding strategies.

@@

\vspace{70 mm}

@@

\vspace{180 mm}
